Simple repression draft

### 2020-06-06 doc/section_04_bayesian_inference.tex
Implemented most of GC's edits plus more of my own

For this model the master equation of interest is
Eq.~\ref{eq:poisson_promoter_cme} with repressor set to zero, i.e.,
\begin{equation}
{d\over dt}p_U(m)(t) = 
        rp_U(m-1)(t) 
        - rp_U(m)(t)
        + (m+1)\gamma p_U(m+1)(t) 
        - \gamma p_U(m)(t),
\end{equation}
whose steady-state solution is given by a Poisson distribution with parameter
$\lambda \equiv r / \gamma$~\cite{Sanchez2013}. The goal of our inference 
problem is then to find the probability distribution for the parameter value
$\lambda$ given the experimental data. By Bayes theorem this can be written as
\begin{equation}
p(\lambda \mid D) = {p(D \mid \lambda) p(\lambda) \over p(D)},
\end{equation}
where $D = \{m_1, m_2, \ldots, m_N \}$ are the single-cell mRNA experimental
counts. As is standard we will neglect the denominator $p(D)$ on the right
hand side since it is independent of $\lambda$ and serves only as a
normalization factor.

@@

As for the Poisson promoter model, this expression for the steady-state mRNA
distribution is exactly the likelihood we want to use when stating Bayes
theorem. Again denoting the single-cell mRNA count data as $D=\{m_1, m_2,\dots,
m_N\}$, here Bayes theorem takes the form
\begin{equation}
p(k_i, b \mid D) \propto p(D\mid k_i,b)p(k_i, b).
\end{equation}
We already have our likelihood -- the product of $N$ negative binomials as
Eq.~\ref{eq:neg_bionom} -- so we only need to choose priors on $k_i$ and $b$.
For the datasets from~\cite{Jones2014} that we are analyzing, as for the Poisson
promoter model above we are still data-rich so the prior's influence remains
weak, but not nearly as weak because the dimensionality of our model has
increased from one parameter to two. Details on the arguments behind our prior
distribution selection are left for Appendix~\ref{sec:bayesian}. We state here
that the natural scale to explore these parameters is logarithmic. This is
commonly the case for parameters for which our previous knowledge based on our
domain expertise spans several orders of magnitude. For this we chose log-Normal
distributions for both $k_i$ and $b$. Details on the mean and variance of these
distributions can be found in Appendix~\ref{sec:bayesian}.

@@

Having found the model's posterior to be well-identified as with the Poisson
promoter, the next step is to compare both models with experimental data. To do
this for the case of the bursty promoter, for each of the parameter samples
shown in Figure~\ref{fig:constit_post_full}(A) we generated negative
bionomial-distributed mRNA counts. As MCMC samples parameter space
proportionally to the posterior distribution, this set of random samples span
the range of possible values that we would expect given the correspondence
between our theoretical model and the experimental data. A similar procedure can
be applied to the Poisson promoter. To compare so many samples with the actual
observed data, we can use empirical cumulative distribution functions (ECDF) of
the distribution quantiles. This representation is shown in
Figure~\ref{fig:constit_post_full}(B). In this example, the median for each
possible mRNA count for the Poisson distribution is shown as a dark green line,
while the lighter green contains 95\% of the randomly generated samples. This
way of representing the fit of the model to the data gives us a sense of the
range of data we might consider plausible, under the assumption that the model
is true. For this case, as we expected given our premise of the Poisson promoter
being wrong, it is quite obvious that the observed data, plotted in black is not
consistent with the Poisson promoter model. An equivalent plot for the bursty
promoter model is shown in blue. Again the darker tone shows the median, while
the lighter color encompasses 95\% of the randomly generated samples. Unlike the
Poisson promoter model, the experimental ECDF closely tracks the posterior
predictive ECDF, indicating this model is actually able to generate the observed
data and increasing our confidence that this model is at least not wrong.

\textit{lacUV5} is our primary target here, since it forms the core of all the
simple repression constructs of~\cite{Jones2014} that we consider in
Section~\ref{sec:rep_kinetics_inference}. Nevertheless, we thought it wise to
apply our bursty promoter model to the other 17 unregulated promoters available
in the single-cell mRNA count dataset from~\cite{Jones2014} as a test that the
model is capturing the essential phenomenology. If the model fit well to all the
different promoters, this would increase our confidence that it would serve well
as a foundation for inferring repressor kinetics later in
Section~\ref{sec:rep_kinetics_inference}. Conversely, were the model to fail on
more than a couple of the other promoters, it would give us pause.

@@

One interesting feature from \fig{fig:constit_post_full}(C) is that burst rate
varies far more widely, over a range of $\sim10^2$, than burst size, confined to
a range of $\lesssim10^1$ (and with the exception of promoter 6, just a span of
3-5x). This suggests that $k_i$, not $b$, is the key dynamic variable that
promoter sequence tunes.

@@

While the associations represented by Eq.~\ref{eq:bursty_equil_corresp1} and
Eq.~\ref{eq:bursty_equil_corresp2} appear to be born out by the data in
Figure~\ref{fig:constit_post_full}, we do not find the association of parameters
they imply to be intuitive. We are also cautious to ascribe too much physical
reality to the parameters. Indeed, part of our point in comparing the various
constitutive promoter models is to demonstrate that these models each provide an
internally self-consistent framework that adequately describes the data, but
attempting to translate between models reveals the dubious physical
interpretation of their parameters.

@@

\begin{figure}%[h!]
\centering
\includegraphics[width=\textwidth]{../figures/main/fig04.pdf}
\caption{\textbf{Simple repression parameter inference and comparison.}
(A) Contours which enclose 50\% and 95\% of the posterior probability mass are
shown for each of several 2D slices of the 9D posterior distribution. The model
assumes one unbinding rate for each operator (Oid, O1, O2) and one binding rate
for each aTc induction concentration (corresponding to an unknown mean repressor
copy number). (B, upper panel) Ratios of our inferred unbinding rates are
compared with operator binding energy differences measured by Garcia and
Phillips~\cite{Garcia2011a} (triangles) and Razo-Mejia et.\
al.~\cite{Razo-Mejia2018} (circles). Blue glyphs compare O2-O1, while orange
compare O1-Oid. (B, lower panel) Unbinding rates for O1 (cyan) and Oid (red)
inferred in this work are compared with single-molecule measurements of the same
from Hammar et.\ at.\ 2014. The comparison requires an assumption about mRNA
lifetime $\gamma^{-1}$, which was not measured, and we plot representative
choices assuming $\gamma^{-1}=3$~min (triangles) or $\gamma^{-1}=5$~min
(squares). (C) Theory-experiment comparison are shown for each of the datasets
used in the inference of the model in (A). Observed single-molecule mRNA counts
data from~\cite{Jones2014} is plotted as black lines. The median of the randomly
generated samples for each condition is plotted as a dark colored line. Lighter
colored bands enclose 95\% of all samples for a given operator/repressor copy
number pair. Samples are sorted by operator and plotted separately for visual
clarity. The unregulated promoter, UV5, is shown with each as a reference
point.}
\label{fig4:repressed_post_full}
\end{figure}

@@

Formally now, denote the set of seven repressor rates to be inferred as
\begin{equation}
\vect{k} =\{k_{Oid}^-, k_{O1}^-, k_{O2}^-,
k_{0.5}^+, k_{1}^+, k_{2}^+, k_{10}^+\},
\end{equation}
where subscripts for dissociation rates $k^-$  indicate the different repressor
binding sites, and subscripts for association rates $k^+$ indicate the
concentration of the small-molecule that controlled the expression of the LacI
repressor (see Appendix~\ref{sec:bayesian}). This is because for this particular
dataset the repressor copy numbers were not measured directly, but it is safe to
assume that a given concentration of the inducer resulted in a specific mean
repressor copy number\cite{Chure2019a}. Also note that the authors
of~\cite{Jones2014} report estimates of LacI copy number per cell rather than
direct measurements. However, these estimates were made assuming the validity of
the equilibrium models in Figure~\ref{fig1:means_cartoons}, and since testing
these models is our present goal, it would be circular logic if we were to make
the same assumption. Therefore we will make no assumptions about the LacI copy
number for a given inducer concentrations.

Having stated the problem, Bayes theorem reads simply
\begin{equation}
p(\vect{k}, k_i, b \mid D)
\propto
p(D \mid\vect{k}, k_i, b) p(\vect{k}, k_i, b),
\end{equation}
where $D$ is again the set of all $N$ observed single-cell mRNA counts
across the various conditions. We assume that individual single-cell
measurements are independent so that the likelihood factorizes as
\begin{equation}
p(D \mid\vect{k}, k_i, b)
= \prod_{j=1}^N p(m\mid \vect{k}, k_i, b)
= \prod_{j=1}^N p(m\mid k_j^+, k_j^-, k_i, b)
\end{equation}
where $k_j^\pm$ represent the appropriate binding and unbinding
rates out of $\vect{k}$ for the $j$-th measured cell. The probability
$p(m\mid k_j^+, k_j^-, k_i, b)$ appearing in the last expression
is exactly Eq.~\ref{eq:p_m_bursty+rep_appdx}, the steady-state
distribution for our bursty model with repression derived in
Section~\ref{sec:gen_fcn_appdx}, which for completeness we reproduce here as
\begin{equation}
\begin{split}
p(m \mid k_R^+, k_R^-, k_i, b) = & ~\frac{
        \Gamma(\alpha + m)\Gamma(\beta + m)\Gamma(k_R^+ + k_R^-)
        }
        {
        \Gamma(\alpha)\Gamma(\beta)\Gamma(k_R^+ + k_R^- + m)
        }
\frac{b^m}{m!}
\\
&\times {_2F_1}(\alpha+m, \beta+m, k_R^++k_R^-+m; -b).
\end{split}
\label{eq:p_m_bursty+rep}
\end{equation}
where $_2F_1$ is the confluent hypergeometric function of the second kind and
$\alpha$ and $\beta$, defined for notational convenience, are
\begin{align}
\begin{split}
\alpha &= \frac{1}{2}
\left(k_i+k_R^-+k_R^+ + \sqrt{(k_i+k_R^-+k_R^+)^2 - 4k_i k_R^-}\right)
\\
\beta &= \frac{1}{2}
\left(k_i+k_R^-+k_R^+ - \sqrt{(k_i+k_R^-+k_R^+)^2 - 4k_i k_R^-}\right).
\end{split}
\end{align}

@@

Next we specify priors. As for the constitutive model, weakly informative
lognormal priors are a natural choice for all our rates. We found that if the
priors were too weak, our MCMC sampler would often become stuck in regions of
parameter space with very low probability density, unable to move. We struck a
balance in choosing our prior widths between helping the sampler run while
simultaneously verifying that the marginal posteriors for each parameter were
not artificially constrained or distorted by the presence of the prior. The only
exception to this is the highly informative priors we placed on $k_i$ and $b$,
since we have strong knowledge of them from our inference of constitutive
promoters above. All details for our prior distributions are listed in 
Appendix~\ref{sec:bayesian}.

@@

Figure~\ref{fig4:repressed_post_full}(B) shows both comparisons, with the top
panel comparing to equilibrium binding energies and the bottom panel comparing
to single-molecule measurements. First consider the top panel and its comparison
between repressor kinetic rates and binding energies. As described in
section~\ref{section_02_means}, if the equilibrium binding energies
from~\cite{Garcia2011a} and~\cite{Razo-Mejia2018} indeed are the physically real
binding energies we believe them to be, then they should be related to the
repressor kinetic rates via Eq.~\ref{eq:deltaFR_eq_noneq_equiv}, which we
restate here,
\begin{equation}
\Delta F_R = \beta\Delta\varepsilon_R - \log(R/N_{NS})
        = - \log(k_R^+/k_R^-).
\label{eq:deltaFR_eq_noneq_equiv_repeat}
\end{equation}
Assuming mass action kinetics implies that $k_R^+$ is proportional to repressor
copy number $R$, or more precisely, it can be thought of as repressor copy
number times some intrinsic per molecule association rate. But since $R$ is not
directly known for our data from~\cite{Jones2014}, we cannot use this equation
directly. Instead we can consider two different repressor binding sites and
compute the \textit{difference} in binding energy between them, since this
difference depends only on the unbinding rates and not on the binding rates.
This can be seen by evaluating Eq.~\ref{eq:deltaFR_eq_noneq_equiv_repeat} for
two different repressor binding sites, labeled (1) and (2), but with the same
repressor copy number $R$, and taking the difference to find
\begin{equation}
\Delta F_R^{(1)} - \Delta F_R^{(2)}
= \beta\Delta\varepsilon_1 - \beta\Delta\varepsilon_2
= - \log(k_R^+/k_1^-) + \log(k_R^+/k_2^-),
\end{equation}
or simply
\begin{equation}
\beta\Delta\varepsilon_1 - \beta\Delta\varepsilon_2
= \log(k_2^-/k_1^-).
\end{equation}
The left and right hand sides of this equation are exactly the horizontal and
vertical axes of the top panel of Figure~\ref{fig4:repressed_post_full}. Since
we inferred rates for three repressor binding sites (O1, O2, and Oid), there are
only two independent differences that can be constructed, and we arbitrarily
chose to plot O2-O1 and O1-Oid in Figure~\ref{fig4:repressed_post_full}(B).
Numerically, we compute values of $k_{O1}^- / k_{Oid}^-$ and $k_{O2}^- /
k_{O1}^-$ directly from our full posterior samples, which conveniently provides
uncertainties as well, as detailed in Appendix~\ref{sec:bayesian}.
\mmnote{Explain in appendix how marginalization with sampling makes this
``trivial'' to compute from samples of the joint posterior, b/c sampling makes
marginalization trivial.} We then compare these log ratios of rates to the
binding energy differences $\Delta\varepsilon_{O1} - \Delta\varepsilon_{Oid}$
and from $\Delta\varepsilon_{O2} - \Delta\varepsilon_{O1}$ as computed from the
values from both~\cite{Garcia2011a} and~\cite{Razo-Mejia2018}. Three of the four
values are within $\sim0.5 k_BT$ of the diagonal representing perfect agreement,
which is comparable to the $\sim\text{few}\times0.1 k_BT$ variability between
the independent determinations of the same quantities between~\cite{Garcia2011a}
and~\cite{Razo-Mejia2018}. The only outlier involves Oid measurements
from~\cite{Razo-Mejia2018}, and as the authors of~\cite{Razo-Mejia2018} note,
this is a difficult measurement of low fluorescence signal against high
background since Oid represses so strongly. We are therefore inclined to regard
the failure of this point to fall near the diagonal as a testament to the
difficulty of the measurement and not as a failure of our theory.

@@

Our broad brush takeaway from Figure~\ref{fig4:repressed_post_full}(C) is
overall of good agreement. There some oddities, in particular the long tails in
the data for Oid, 1~ng/mL, and O2, 0.5~ng/mL. The latter is especially odd since
it extends beyond the tail of the unregulated UV5 distribution. This is a
relatively small number of cells, however, so whether this is a peculiarity of
the experimental data, a statistical fluke of small numbers, or a real
biological effect is unclear. It is conceivable that there is some very slow
timescale switching dynamics that could cause this bimodality, although it is
unclear why it would only appear for specific repressor copy numbers. There is
also a small offset between experiment and simulation for O2 at the higher
repressor copy numbers, especially at 2 and 10~ng/mL.
From the estimate of repressor copy numbers from~\cite{Jones2014}, it is
possible that the repressor copy numbers here are becoming large enough to
partially invalidate our assumption of a separation of timescales between
burst duration and repressor association rate. Another possibility is that
the very large number of zero mRNA counts for Oid, 2~ng/mL is skewing it's partner
datasets through the shared association rate.
None of these fairly minor quibbles cause us to
seriously doubt the overall correctness of our model, which further validates
its use to compare the equilibrium models' binding energies to the
nonequilibrium models' repressor kinetics, as we originally set out to do.

>>>>

For this model the master equation of interest is
Eq.~\ref{eq:poisson_promoter_cme} with repressor set to zero, i.e.,
\begin{equation}
{d\over dt}p_U(m)(t) = 
        rp_U(m-1)(t) 
        - rp_U(m)(t)
        + (m+1)\gamma p_U(m+1)(t) 
        - \gamma p_U(m)(t),
\end{equation}
whose steady-state solution is given by a Poisson distribution with parameter
$\lambda \equiv r / \gamma$~\cite{Sanchez2013}. The goal of our inference 
problem is then to find the probability distribution for the parameter value
$\lambda$ given the experimental data. By Bayes' theorem this can be written as
\begin{equation}
p(\lambda \mid D) = {p(D \mid \lambda) p(\lambda) \over p(D)},
\end{equation}
where $D = \{m_1, m_2, \ldots, m_N \}$ are the single-cell mRNA experimental
counts. As is standard we will neglect the denominator $p(D)$ on the right
hand side since it is independent of $\lambda$ and serves only as a
normalization factor.

@@

As for the Poisson promoter model, this expression for the steady-state mRNA
distribution is exactly the likelihood we want to use when stating Bayes
theorem. Again denoting the single-cell mRNA count data as $D=\{m_1, m_2,\dots,
m_N\}$, here Bayes' theorem takes the form
\begin{equation}
p(k_i, b \mid D) \propto p(D\mid k_i,b)p(k_i, b).
\end{equation}
We already have our likelihood -- the product of $N$ negative binomials as
Eq.~\ref{eq:neg_bionom} -- so we only need to choose priors on $k_i$ and $b$.
For the datasets from~\cite{Jones2014} that we are analyzing, as for the Poisson
promoter model above we are still data-rich so the prior's influence remains
weak, but not nearly as weak because the dimensionality of our model has
increased from one parameter to two. Details on the arguments behind our prior
distribution selection are left for Appendix~\ref{sec:bayesian}. We state here
that the natural scale to explore these parameters is logarithmic. This is
commonly the case for parameters for which our previous knowledge based on our
domain expertise spans several orders of magnitude. For this we chose log-normal
distributions for both $k_i$ and $b$. Details on the mean and variance of these
distributions can be found in Appendix~\ref{sec:bayesian}.

@@

Having found the model's posterior to be well-identified as with the Poisson
promoter, the next step is to compare both models with experimental data. To do
this for the case of the bursty promoter, for each of the parameter samples
shown in Figure~\ref{fig:constit_post_full}(A) we generated negative
bionomial-distributed mRNA counts. As MCMC samples parameter space
proportionally to the posterior distribution, this set of random samples span
the range of possible values that we would expect given the correspondence
between our theoretical model and the experimental data. A similar procedure can
be applied to the Poisson promoter. To compare so many samples with the actual
observed data, we can use empirical cumulative distribution functions (ECDF) of
the distribution quantiles. This representation is shown in
Figure~\ref{fig:constit_post_full}(B). In this example, the median for each
possible mRNA count for the Poisson distribution is shown as a dark green line,
while the lighter green contains 95\% of the randomly generated samples. This
way of representing the fit of the model to the data gives us a sense of the
range of data we might consider plausible, under the assumption that the model
is true. For this case, as we expected given our premise of the Poisson promoter
being wrong, it is quite obvious that the observed data, plotted in black is not
consistent with the Poisson promoter model. An equivalent plot for the bursty
promoter model is shown in blue. Again the darker tone shows the median, while
the lighter color encompasses 95\% of the randomly generated samples. Unlike the
Poisson promoter model, the experimental ECDF closely tracks the posterior
predictive ECDF, indicating this model is actually able to generate the observed
data and increasing our confidence that this model
is sufficient to parametrize the physical reality of the system.

The commonly used promoter sequence \textit{lacUV5} is our primary
target here, since it forms the core of all the
simple repression constructs of~\cite{Jones2014} that we consider in
Section~\ref{sec:rep_kinetics_inference}. Nevertheless, we thought it wise to
apply our bursty promoter model to the other 17 unregulated promoters available
in the single-cell mRNA count dataset from~\cite{Jones2014} as a test that the
model is capturing the essential phenomenology. If the model fit well to all the
different promoters, this would increase our confidence that it would serve well
as a foundation for inferring repressor kinetics later in
Section~\ref{sec:rep_kinetics_inference}. Conversely, were the model to fail on
more than a couple of the other promoters, it would give us pause.

@@

One interesting feature from \fig{fig:constit_post_full}(C) is that burst rate
varies far more widely, over a range of $\sim10^2$, than burst size, confined to
a range of $\lesssim10^1$ (and with the exception of promoter 6, just a span of
3 to 5-fold). This suggests that $k_i$, not $b$, is the key dynamic variable that
promoter sequence tunes.

@@

While the associations represented by Eq.~\ref{eq:bursty_equil_corresp1} and
Eq.~\ref{eq:bursty_equil_corresp2} appear to be borne out by the data in
Figure~\ref{fig:constit_post_full}, we do not find the association of parameters
they imply to be intuitive. We are also cautious to ascribe too much physical
reality to the parameters. Indeed, part of our point in comparing the various
constitutive promoter models is to demonstrate that these models each provide an
internally self-consistent framework that adequately describes the data, but
attempting to translate between models reveals the dubious physical
interpretation of their parameters.

@@

\begin{figure}%[h!]
\centering
\includegraphics[width=\textwidth]{../figures/main/fig04.pdf}
\caption{\textbf{Simple repression parameter inference and comparison.}
(A) Contours which enclose 50\% and 95\% of the posterior probability mass are
shown for each of several 2D slices of the 9D posterior distribution. The model
assumes one unbinding rate for each operator (Oid, O1, O2) and one binding rate
for each aTc induction concentration (corresponding to an unknown mean repressor
copy number).
(B, upper) Ratios of our inferred unbinding rates are
compared with operator binding energy differences measured by Garcia and
Phillips~\cite{Garcia2011a} (triangles) and Razo-Mejia et.\
al.~\cite{Razo-Mejia2018} (circles). Blue glyphs compare O2-O1, while orange
compare O1-Oid. Points with perfect agreement would lie on the dotted line.
(B, lower) Unbinding rates for O1 (cyan) and Oid (red)
inferred in this work are compared with single-molecule measurements
from Hammar et.\ al.~\cite{Hammar2014}. We plot the comparison assuming illustrative mRNA
lifetimes of $\gamma^{-1}=3$~min (triangles) or $\gamma^{-1}=5$~min
(squares). Dotted line is as in upper panel.
(C) Theory-experiment comparison are shown for each of the datasets
used in the inference of the model in (A). Observed single-molecule mRNA counts
data from~\cite{Jones2014} are plotted as black lines. The median of the randomly
generated samples for each condition is plotted as a dark colored line. Lighter
colored bands enclose 95\% of all samples for a given operator/repressor copy
number pair.
%Samples are sorted by operator and plotted separately for visual clarity.
The unregulated promoter, \textit{lacUV5}, is shown with each as a reference.}
\label{fig4:repressed_post_full}
\end{figure}

@@

Formally now, denote the set of seven repressor rates to be inferred as
\begin{equation}
\vect{k} =\{k_{Oid}^-, k_{O1}^-, k_{O2}^-,
k_{0.5}^+, k_{1}^+, k_{2}^+, k_{10}^+\},
\end{equation}
where subscripts for dissociation rates $k^-$  indicate the different repressor
binding sites, and subscripts for association rates $k^+$ indicate the
concentration of the small-molecule that controlled the expression of the LacI
repressor (see Appendix~\ref{sec:bayesian}). This is because for this particular
dataset the repressor copy numbers were not measured directly, but it is safe to
assume that a given concentration of the inducer resulted in a specific mean
repressor copy number~\cite{Chure2019a}. Also note that the authors
of~\cite{Jones2014} report estimates of LacI copy number per cell rather than
direct measurements. However, these estimates were made assuming the validity of
the equilibrium models in Figure~\ref{fig1:means_cartoons}, and since testing
these models is our present goal, it would be circular logic if we were to make
the same assumption. Therefore we will make no assumptions about the LacI copy
number for a given inducer concentrations.

Having stated the problem, Bayes' theorem reads
\begin{equation}
p(\vect{k}, k_i, b \mid D)
\propto
p(D \mid\vect{k}, k_i, b) p(\vect{k}, k_i, b),
\end{equation}
where $D$ is again the set of all $N$ observed single-cell mRNA counts
across the various conditions. We assume that individual single-cell
measurements are independent so that the likelihood factorizes as
\begin{equation}
p(D \mid\vect{k}, k_i, b)
= \prod_{j=1}^N p(m\mid \vect{k}, k_i, b)
= \prod_{j=1}^N p(m\mid k_j^+, k_j^-, k_i, b)
\end{equation}
where $k_j^\pm$ represent the appropriate binding and unbinding
rates out of $\vect{k}$ for the $j$-th measured cell. The probability
$p(m\mid k_j^+, k_j^-, k_i, b)$ appearing in the last expression
is exactly Eq.~\ref{eq:p_m_bursty+rep_appdx}, the steady-state
distribution for our bursty model with repression derived in
Section~\ref{sec:gen_fcn_appdx}, which for completeness we reproduce here as
\begin{equation}
\begin{split}
p(m \mid k_R^+, k_R^-, k_i, b) = & ~\frac{
        \Gamma(\alpha + m)\Gamma(\beta + m)\Gamma(k_R^+ + k_R^-)
        }
        {
        \Gamma(\alpha)\Gamma(\beta)\Gamma(k_R^+ + k_R^- + m)
        }
\frac{b^m}{m!}
\\
&\times {_2F_1}(\alpha+m, \beta+m, k_R^++k_R^-+m; -b).
\end{split}
\label{eq:p_m_bursty+rep}
\end{equation}
where $_2F_1$ is the confluent hypergeometric function of the second kind and
$\alpha$ and $\beta$, defined for notational convenience, are
\begin{align}
\begin{split}
\alpha &= \frac{1}{2}
\left(k_i+k_R^-+k_R^+ + \sqrt{(k_i+k_R^-+k_R^+)^2 - 4k_i k_R^-}\right)
\\
\beta &= \frac{1}{2}
\left(k_i+k_R^-+k_R^+ - \sqrt{(k_i+k_R^-+k_R^+)^2 - 4k_i k_R^-}\right).
\end{split}
\end{align}

@@

Next we specify priors. As for the constitutive model, weakly informative
log-normal priors are a natural choice for all our rates. We found that if the
priors were too weak, our MCMC sampler would often become stuck in regions of
parameter space with very low probability density, unable to move. We struck a
balance in choosing our prior widths between helping the sampler run while
simultaneously verifying that the marginal posteriors for each parameter were
not artificially constrained or distorted by the presence of the prior.
% The only
% exception to this is the highly informative priors we placed on $k_i$ and $b$,
% since we have strong knowledge of them from our inference of constitutive
% promoters above.
All details for our prior distributions are listed in 
Appendix~\ref{sec:bayesian}.

@@

Figure~\ref{fig4:repressed_post_full}(B) shows both comparisons, with the top
panel comparing to equilibrium binding energies and the bottom panel comparing
to single-molecule measurements. First consider the top panel and its comparison
between repressor kinetic rates and binding energies. As described in
section~\ref{section_02_means}, if the equilibrium binding energies
from~\cite{Garcia2011a} and~\cite{Razo-Mejia2018} indeed are the physically real
binding energies we believe them to be, then they should be related to the
repressor kinetic rates via Eq.~\ref{eq:deltaFR_eq_noneq_equiv}, which we
restate here,
\begin{equation}
\Delta F_R = \beta\Delta\varepsilon_R - \log(R/N_{NS})
        = - \log(k_R^+/k_R^-).
\label{eq:deltaFR_eq_noneq_equiv_repeat}
\end{equation}
Assuming mass action kinetics implies that $k_R^+$ is proportional to repressor
copy number $R$, or more precisely, it can be thought of as repressor copy
number times some intrinsic per molecule association rate. But since $R$ is not
directly known for our data from~\cite{Jones2014}, we cannot use this equation
directly. Instead we can consider two different repressor binding sites and
compute the \textit{difference} in binding energy between them, since this
difference depends only on the unbinding rates and not on the binding rates.
This can be seen by evaluating Eq.~\ref{eq:deltaFR_eq_noneq_equiv_repeat} for
two different repressor binding sites, labeled (1) and (2), but with the same
repressor copy number $R$, and taking the difference to find
\begin{equation}
\Delta F_R^{(1)} - \Delta F_R^{(2)}
= \beta\Delta\varepsilon_1 - \beta\Delta\varepsilon_2
= - \log(k_R^+/k_1^-) + \log(k_R^+/k_2^-),
\end{equation}
or simply
\begin{equation}
\beta\Delta\varepsilon_1 - \beta\Delta\varepsilon_2
= \log(k_2^-/k_1^-).
\end{equation}
The left and right hand sides of this equation are exactly the horizontal and
vertical axes of the top panel of Figure~\ref{fig4:repressed_post_full}. Since
we inferred rates for three repressor binding sites (O1, O2, and Oid), there are
only two independent differences that can be constructed, and we arbitrarily
chose to plot O2-O1 and O1-Oid in Figure~\ref{fig4:repressed_post_full}(B).
Numerically, we compute values of $k_{O1}^- / k_{Oid}^-$ and $k_{O2}^- /
k_{O1}^-$ directly from our full posterior samples, which conveniently provides
uncertainties as well, as detailed in Appendix~\ref{sec:bayesian}.
% \mmnote{Explain in appendix how marginalization with sampling makes this
% ``trivial'' to compute from samples of the joint posterior, b/c sampling makes
% marginalization trivial.}
We then compare these log ratios of rates to the
binding energy differences $\Delta\varepsilon_{O1} - \Delta\varepsilon_{Oid}$
and from $\Delta\varepsilon_{O2} - \Delta\varepsilon_{O1}$ as computed from the
values from both~\cite{Garcia2011a} and~\cite{Razo-Mejia2018}. Three of the four
values are within $\sim0.5~k_BT$ of the diagonal representing perfect agreement,
which is comparable to the $\sim\text{few}\times0.1~k_BT$ variability between
the independent determinations of the same quantities between~\cite{Garcia2011a}
and~\cite{Razo-Mejia2018}. The only outlier involves Oid measurements
from~\cite{Razo-Mejia2018}, and as the authors of~\cite{Razo-Mejia2018} note,
this is a difficult measurement of low fluorescence signal against high
background since Oid represses so strongly. We are therefore inclined to regard
the failure of this point to fall near the diagonal as a testament to the
difficulty of the measurement and not as a failure of our theory.

@@

Our broad brush takeaway from Figure~\ref{fig4:repressed_post_full}(C) is
overall of good agreement. There some oddities, in particular the long tails in
the data for Oid, 1~ng/mL, and O2, 0.5~ng/mL. The latter is especially odd since
it extends beyond the tail of the unregulated UV5 distribution. This is a
relatively small number of cells, however, so whether this is a peculiarity of
the experimental data, a statistical fluke of small numbers, or a real
biological effect is unclear. It is conceivable that there is some very slow
timescale switching dynamics that could cause this bimodality, although it is
unclear why it would only appear for specific repressor copy numbers. There is
also a small offset between experiment and simulation for O2 at the higher
repressor copy numbers, especially at 2 and 10~ng/mL.
From the estimate of repressor copy numbers from~\cite{Jones2014}, it is
possible that the repressor copy numbers here are becoming large enough to
partially invalidate our assumption of a separation of timescales between
burst duration and repressor association rate. Another possibility is that
the very large number of zero mRNA counts for Oid, 2~ng/mL is skewing its partner
datasets through the shared association rate.
None of these fairly minor quibbles cause us to
seriously doubt the overall correctness of our model, which further validates
its use to compare the equilibrium models' binding energies to the
nonequilibrium models' repressor kinetics, as we originally set out to do.
